---
title: "Functions"
description: |
  Repeating things? Let's functionalize it!
output: 
  distill::distill_article:
    toc: true
    toc_float: true
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
```

When we see code being repeated more than once, functions are a great way to reduce duplication. Even if we call a function only once, they can be a nice way to break up large complicated processes.

# What's in a function?

1. The Formals
2. The Body
3. The Environment

To define a function here's the basic skeleton

```{r, eval=FALSE}
my_function_name <- function() {
  
}

```

# Let's create one

Suppose we'd like to do some cleaning to each CHAS table in the same manner. Let's create one that does the following: 

  - filter for PSRC counties
  - pivot longer (so columns that start with 'T' are not across the table)
  - dissect the column containing the former 'T...' headers into three columns
      - create 'table' field
      - create a 'type' field to identify whether values are 'est' or 'moe'
      - create a 'sort' field
      
```{r eval=FALSE}
# Read-in one file as a data frame
file_01 <- read_csv(here('data', '050', 'Table9.csv'))

# define the skeleton of our function
# add table as a parameter
clean_table <- function(table) {
  
}

```

Fill in the body with the argument to clean

```{r eval=FALSE}
clean_table <- function(table) {
  table %>% 
    filter(st == 53 & cnty %in% c('033', '035', '053', '061')) %>% 
    pivot_longer(cols = str_subset(colnames(table), "^T.*"), 
                 names_to = 'header', 
                 values_to = 'value') %>% 
    mutate(table = str_extract(header, "^T\\d*(?=_)"), # string starting with T and numeric digits followed by _
           type = str_extract(header, "(?<=_)\\w{3}"), # 3 letters preceded by _
           sort = str_extract(header, "\\d+$")) # last numeric digits at the end of the string
}
```

<aside>
Functions will generally return the last evaluated expression. With the piping (`%>%`) in dplyr, our example is essentially a one liner expression. You can always add `return(<name of object>)` to explicitly return a specific object whenever your function is called.
</aside>

## Call the function

```{r eval=FALSE}
t9 <- clean_table(file_01)
```

Try with other files

```{r eval=FALSE}
file_02 <- read_csv(here('data', '050', 'Table10.csv'))
file_03 <- read_csv(here('data', '050', 'Table11.csv'))

t10 <- clean_table(file_02)
t11 <- clean_table(file_03)
```

If we forgot a step in the cleaning process, we can always edit the function and re-run our script

```{r eval=FALSE}
# Let's make this edit to our function that will convert the sort column from string to numeric
sort = as.numeric(str_extract(header, "\\d+$"))
```

# Benefits of creating functions

- Easier editing of code
- Reduce redundancy
- Break long processes into chunks


