---
title: "Loops"
description: |
  Let the machine do it!

output: 
  distill::distill_article:
    toc: true
    toc_float: true
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We could explicitly call the function as many times need be to clean the tables we're interested in.

```{r eval=FALSE}
table9 <- clean_table(file_01)

table10 <- clean_table(file_02)

table11 <- clean_table(file_03)
```

But that would be redundant! There would be three extra lines of code and three extra variable names in your global environment to keep track of.

# Anatomy

A `for` loop has three parts:

1. The Output: where the stuff will be stored
2. The Sequence: code within `()` that shows what to loop over
3. The Body: code within `{}` that does the work

## Sequence

The sequence lies within the `()`. It will follow this structure:
(<variable name of your choice> in <list or vector>)

The sequence tells the machine what to loop over. The variable name will represent a single element within the list or vector.

In the example below, with every iteration, `df` will be a counter and represent a different data frame in `l`

```{r eval=FALSE}
l <- list(file_01, file_02, file_03)

# for every data frame (df) in list (l)...
for(df in l) {
  
}
```

Try printing the `head()` of each data frame in our list

```{r eval=FALSE}
l <- list(file_01, file_02, file_03) 

for(df in l) {
  print(head(df))
}

```

Try printing a version of each data frame with `clean_table()`

```{r eval=FALSE}
for(df in l) {
  print(clean_table(df))
}
```

## Output
Now instead of printing stuff, let's store stuff in a list!

A way to use both lists and loops is to read in data. Let's create a loop to read-in csvs 1 through 11 and store them into a list.

- Initiate the Output (`dfs`) to store the end result (data frames from csvs)
- Construct the sequence you'll be looping over (`csv`) (file names of csvs)
- Read csv (`t`)
```{r}
dfs <- list()

csv <- paste0('Table', 1:11, '.csv')

for(c in csv) {
  t <- read_csv(here('data', '050', c))
  dfs[[c]] <- t
}
```




## Sequencing alternatives


```{r eval=FALSE}

for(df in 1:length(l)) {
  
}

```

# Alter the Flow

## Break

## Next







